Survival of the Fittest: Variable Selection on Agricultural Data from the Gal\'apagos Islands

### 2018-03-08 Paper/paper_draft.tex
completion of paper draft

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts, amssymb, amsmath}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{csvsimple}
\usepackage{subcaption}

@@

\title{Survival of the Fittest: Variable Selection on Galapagos Agricultural Data}

@@

\maketitle
\begin{abstract}
  This paper...

  \end{abstract}

@@

The Galapagos Islands make for a feasible and significant case study of complex systems. Due to its relative isolation and smaller size the interaction of factors can more realistically be modeled for the Galapagos Islands than other systems. Yet the Galapagos Islands also represents an important example of the competing forces of resource conservation and economic development in a rapidly changing environment. Prior work has created agent-based models of the Galapagos, but with limited interaction parameters between agents, particularly in regards to farm success walsh. In order to create a more detailed, and perhaps more accurate, simulation, the relationships between different factors on the island must be better understood. This work aims to sort through a large number of possible relationships and clarify the empirically most significant ones for future study and incorporation into simulations.

@@

The data available to study the dynamics between agricultural measures and related factors primarily 
comes from the Censo de las Unidades de Producci\'on Agropecuaria (upa) de Gal\'apagos (Census of Agricultural Production 
Units (UPA) of Galapagos) ref census. This is a self-reported survey with data from 755 
farms detailing the production and sale of crops and livestock, agricultural expenses, and land use 
decisions. The response rate for this was ??. In addition to the census, data is also available from ... including information on water, energy and road access.

categories...

In total, under the direction of the client 239 variables were selected for consideration in modeling 
relationships between predictors and five outcome variables of interest. Some of the five outcome variables of interest come 
directly from survey responses, while others were derived from a combination of multiple variables. When modeling derived 
outcome variables all variables used in its calculation were removed from consideration. The client also denoted specific 
predictor variables to exclude from particular models when their inclusion would not be beneficial. For example, while the 
amount of crops sold in pounds was not directly used to calculate net income, the obvious relationship existent precluded it 
from inclusion. In addition, predictor variables that met one or more of the following criteria were removed prior to modeling: 
zero variance, extremely high (>0.99) or perfect correlation with other predictor variables, or linear dependence with other 
predictor variables (that is, two or more predictor variables could be linearly combined to create another predictor variable). The 
exact number of predictor variables included in each model varied slightly, but there were approximately 200 predictors variables examined for each model after the preceding steps were taken.

@@

The remainder of this report is divided into four sections, Section 2: Modeling, Section 3: Results, Section 4: Statistical 
Methods, and Section 5: Limitations and Future Work. Section 2 provides an brief overview of the analysis so that the results can be understood. In Section 3 results for each of the 5 outcome variables is provided, where standard tables and graphs are repeated for each. A more in-depth explanation of the statistical methods used is contained in Section 4, but this section can be referenced as needed. Section 5 details important considerations when interpreting the results and suggests possible avenues for future work. Lastly, References and the Appendix, including additional tables and figures, can be found at the end.

@@

The primary challenge in this analysis is the vast amount of potential predictor variables. This challenge is twofold; 1.) when 
the number of predictors is large the calculation of a reliable model is difficult and 2.) interpreting the coefficients of many 
predictors simultaneously is not an easy task for humans (and will make resulting simulations overly complicated). For this 
reason, the analysis focuses on the use of two variable selection techniques that aim to build a linear model with a subset of 
the available variables that still maintains a strong explanatory/predictive performance.

Secondarily, when performing standard linear regression the error is assumed to be normally 
distributed, which means the outcome variable should be roughly normally distributed. When this is not true, as is the case for 
several outcome variables in this study, a poorly fitting model will be found with unreliable coefficients. In order to address this 
issue transformations to the data and modifications to the standard linear model will be considered.

@@

A brief overview of the statistical methods used is presented here to allow for understanding of results, 
but for further details see Section refstatmethods Statistical Methods. For each of the outcome variables of interest 
we build a set of linear models using the appropriate subset of predictors. Each relationship is modeled 
using Forward Selection and ElasticNet regression. Forward Selection fits a linear model by progressively adding variables to 
the model until a best fit is found. This results in only some of the variables being included, chosen in a discrete manner. 
ElasticNet regression fits a linear model by limiting the size of the coefficients so that they are smaller than in standard least 
squares, and for many variables actually shrunken to zero. Similar to Forward Selection this results in a smaller model, but 
variable selection can be carefully tuned as optimization is done in a more continuous way. In general, these techniques have 
slightly different aims. Forward Selection chooses a model that best explains the variance in the dataset at hand. Elasticnet 
chooses a model that can best make predictions on new data. Depending on the goals of analysis, one technique is not 
necessarily better so we do not compare the two quantitatively, but instead offer both results as varying perspectives on 
variable selection. While a variable being chosen by both methods provides stronger evidence that an important relationship exists, disagreement should suggest exploring both possibilities instead of one method necessarily being incorrect.

@@

The first three outcome variables of interest can be grouped together under the category of farm success; labeled as 
productivity, net income and number of workers supported. Productivity is calculated as the total pounds of crops and livestock 
produced divided by the farm surface area. Net income is calculated as the the earnings from all products 
sold minus total expenses. Number of workers supported is calculated as the total labor expenditures divided by a standard 
full-time worker's salary.

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.8\textwidth]{production_histograms.pdf}
\caption{Histogram of Productivity...}
\label{figure:prod_hist}
\end{figure}

The histograms of the Productivity variable (Figure \ref{figure:prod_hist}) show a strong skewness, both when looking at all 
observations, and when zooming into observations between 0-10,000 lbs/hectare. To achieve a distribution closer to normal 
(bell-curved), which will benefit the linear model, we take the $log_{10}$ transformation with resulting data shown in the last 
plot. Since the log transformation cannot be performed on zero values, we remove the ?? occasions of this from the dataset. 
Beyond the mathematical constraint, farms with zero production perhaps are not farms as typically defined.

\begin{table}[h!]
\centering
\begin{tabular}{lc|cl}\hline%
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection} 
\csvreader[head to column names]{production_top5names.csv}{}%
{\\\elasticnet & \forward}%
\\\hline
\end{tabular}
\caption{Modeling of Productivity, Top 5 features for each of the three methods}
\label{table:prod_top5}
\end{table}

We build linear models using both methods, Elasticnet and Forward Selection, on the log-transformed 
productivity variable, recording an optimal model of any size and the best 5 variable model for each. The size of 5 variables is 
chosen for its interpretability and not for any specific statistical property. The results of the best 5 variable model are shown in 
Table \ref{table:prod_top5}, listed in alphabetical order.  Next to variable names the direction of the relationship is indicated with 
a (+) or (-). There is some overlap between the variables selected by each of the methods, but also unique choices made by 
each method. The Root Mean-Squared Error (RMSE) for the 5 variable model Elasticnet model is 0.78 and the $R^2$ for 
Forward Selection is 0.48. The RMSE is on average how far the predicted value is from the actual value and $R^2$ is the 
percentage of variability in the outcome variable is explained by the model.

Plots from the optimal models for Elasticnet and Forward Selection are shown in Figure \ref{figure:prod_opt}. The cross-validation plot for Elasticnet can be understood as follows: the 
horizontal axis shows the number of variables included in the model (on top) and the corresponding lambda ($\lambda$) value 
(on the bottom), the vertical axis shows the Mean-Squared Error (MSE) represented as the red dots and surrounded by bars 
showing the standard deviation. The vertical dashed line to the left, $\lambda_{min}$, is found at the minimum MSE and the 
vertical dashed line to the right, $\lambda_{1se}$ is at the largest lambda within one standard error of the minimum. The idea 
behind $\lambda_{1se}$ is that similar error performance can be achieved with a smaller model, in this case a model with 40 
less variables. Since our goal is to select a small amount of variables, we will generally use the model found at $\lambda_{1se}
$. For Forward Selection the plot is much more straightforward, we plot the number of variables included versus the 
information criterion that we would like to minimize and mark the optimal point in red.

The convex shape of the plots highlights a common trend in variable selection; not including enough variables does not provide 
enough information, but beyond a certain point adding more variables may not be worth the added complexity. The optimal 
model chosen for Elasticnet includes 45 variables (not counting the intercept term) and using Forward Selection we choose a 
model of 26 variables, almost all of which are also included in the Elasticnet model. The coefficients estimated for both models 
can be found in Table \ref{table:prod_full} in the Appendix. For the full models the RMSE and $R^2$ are 0.61 and 0.63 
respectively. These numbers suggest that while the 5 variable model is helpful, there is a decent amount of information to be 
gained by adding more variables. Diagnostics of the linear fit of the optimal Elasticnet and Forward Selection models (plots shown in Figure \ref{figure:resids_prod} in the Appendix) do not raise any concerns.

\begin{figure}[!tbp]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{elastic_cv_production.pdf}
\caption{Cross-validation plots for Elasticnet...}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{forward_nvars_production.pdf}
\caption{Number of variables Forward Selection plots...}
\end{subfigure}
\caption{Production Variable Selection}
\label{figure:prod_opt}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.75\textwidth]{netincome_histograms.pdf}
\caption{Histogram of Net Income...}
\label{figure:netincome_hist}
\end{figure}

The histograms for Net Income (Figure \ref{figure:netincome_hist}) show a symmetric shape, but a very spiky center and a few 
observations wide in the tails. We attempt to fit a model on the full dataset, but find the observations with large absolute 
values are obscuring other possible information in the model. We then remove the 28 observations that are beyond one 
standard deviation (30,478) from the mean in either direction. The following models are fit on this reduced dataset of 727 
observations.

\begin{table}[h!]
\centering
\begin{tabular}{c|c|c}\hline%
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{netincome_top5names.csv}{}%
{\\\elasticnet & \forward}%
\\\hline
\end{tabular}
\caption{Modeling of Net Income, Top 5 features for each of the three methods}
\label{table:netincome_top5}
\end{table}

The results of the best 5 variable model are shown in \ref{table:netincome_top5}, with a RMSE of 5790.35 for Elasticnet and a $R^2$ of 0.15 for Forward Selection. For interpretation of the RMSE it is important to keep in mind the scale for Net Income is much larger than that of log productivity.

\begin{figure}[!tbp]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{elastic_cv_netincome.pdf}
\caption{Cross-validation plots for Elasticnet...}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{forward_nvars_netincome.pdf}
\caption{Number of variables Forward Selection plots...}
\end{subfigure}
\caption{Net Income Variable Selection}
\label{figure:netincome_opt}
\end{figure}

The cross-validation plot for Net Income show wider error bars throughout the range of model sizes and just using the average 
Net Income would predict nearly as well as any other model. Since here $\lambda_{1se}$ only includes the intercept, we 
choose the optimal Elasticnet model to be at $\lambda_{min}$, which includes 9 variables. For Forward Selection, we also 
include 9 variables, with all but two overlapping with the Elasticnet choices. The coefficents estimated for both models can be 
found in Table \ref{table:netincome_full} in the Appendix. The full models had an RMSE and $R^2$ of 5768.30 and 0.19, 
respectively. Since the full models are not much larger than the 5 variable models, the small increases are not surprising. Diagnostics of the linear fit of the optimal Elasticnet and Forward Selection models (plots shown in Figure 
\ref{figure:resids_netincome} in the Appendix) do not follow assumptions as closely as for the Production models, but are not so 
concerning as to disqualify the results. On the whole, the results from the various plots and diagnostics suggest that the relationships found for Net Income are worth investigating, but that a linear relationship is not very strong.

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.75\textwidth]{worker_histograms.pdf}
\caption{Histogram of Workers...}
\label{figure:workers_hist}
\end{figure}

The first plot in Figure \ref{figure:workers_hist} shows that a large percentage of the farms are not able to support any workers. 
For this reason, we divide up the modeling task for this outcome variable. First, we use logistic regression to model the binary 
variable of whether a farm supports more than zero or zero workers. Secondly, we use linear regression to model the quantity 
of workers for just those 264 farms with a positive number of workers supported.

\begin{table}[h]
\centering
\begin{tabular}{c|c|c}\hline%
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{workers_binary_top5names.csv}{}%
{\\\elasticnet & \forward}%
\\\hline
\end{tabular}
\caption{Modeling of Binary Workers , Top 5 features for each method}
\label{table:workers_binary_top5}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{c|c|c}\hline%
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{workers_nonzero_top5names.csv}{}%
{\\\elasticnet & \forward}%
\\\hline
\end{tabular}
\caption{Modeling of Nonzero Workers , Top 5 features for each method}
\label{table:workers_nonzero_top5}
\end{table}

Sometimes Elasticnet will simultaneously choose to include multiple variables at the same time, in this case there is not a 5 
variable model so we show the 6 variable model for Elasticnet. For logistic regression we can measure performance with the 
misclassification rate, which on average is 0.31 on the test set for Elasticnet and 0.26 on the full dataset for Forward Selection. 
For the linear model 5 variable model the RMSE is 0.46 and the $R^2$  0.30. Several variables are found to be most helpful for both the logistic regression and linear regression models, but there is still a fair bit of difference.

\begin{figure}[!tbp]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{elastic_cv_workers_binary.pdf}
\caption{Cross-validation plots for Elasticnet...}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{forward_nvars_workers_binary.pdf}
\caption{Number of variables Forward Selection plots...}
\end{subfigure}
\caption{Workers Binary Variable Selection}
\label{figure:workers_binary_opt}
\end{figure}

\begin{figure}[!tbp]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{elastic_cv_workers_nonzero.pdf}
\caption{Cross-validation plots for Elasticnet...}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{forward_nvars_workers_nonzero.pdf}
\caption{Number of variables Forward Selection plots...}
\end{subfigure}
\caption{Workers Positive Variable Selection}
\label{figure:workers_nonzero_opt}
\end{figure}

For logistic regression Elasticnet chooses a model of size 6 and Forward Selection chooses a model of size 7. Since these 
models are not much larger the misclassification rates remain at 0.31 and 0.26, respectively. For linear regression, the full 
models for Elasticnet and Forward Selection have a RMSE and $R^2$ of 0.46 and 0.34, respectively. The Elasticnet full model 
has 5 variables and the Forward Selection full model has 8 variables.  In the Appendix, the coefficients estimated for logistic 
regression models can be found in Table \ref{table:workers_binary_full} and the coefficients estimated for linear regression 
models can be found in Table \ref{table:workers_nonzero_full}.

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.75\textwidth]{invasive_histograms.pdf}
\caption{Histogram of Invasive...}
\label{figure:invasive_hist}
\end{figure}

\begin{table}[h!]
\centering
\begin{tabular}{c|c|c}\hline%
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{invasive_binary_top5names.csv}{}%
{\\\elasticnet & \forward}%
\\\hline
\end{tabular}
\caption{Modeling of Binary Invasive , Top 5 features for each method}
\label{table:invasive_binary_top5}
\end{table}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts, amssymb, amsmath}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{csvsimple}
\usepackage{subcaption}
\usepackage{numprint}

@@

\title{Survival of the Fittest: Variable Selection on Agricultural Data from the Gal\'apagos Islands}

@@

\maketitle
\begin{abstract}
  Variable selection is an important first step when analyzing datasets with a large number of potential predictor variables. We 
  apply two techniques, Forward Selection and Elasticnet,  to find the most important factors in a dataset detailing over 200 
  socioeconomic measurements for 755 farms on the Gal\'apagos Islands. Modeling five different outcome variables, we find 
  the data available has the strongest linear relationships with Productivity and Land Use Choices. For each of the outcome 
  variables we present the top five predictor variables as well as a full set of coefficients for the optimally predictive model.
  \end{abstract}

@@

The Gal\'apagos Islands make for a feasible and significant case study of complex systems. Due to its relative isolation and smaller size the interaction of factors can more realistically be modeled for the Gal\'apagos Islands than other systems. Yet the Gal\'apagos Islands also represents an important example of the competing forces of resource conservation and economic development in a rapidly changing environment. Prior work has created agent-based models of the Gal\'apagos, but with limited interaction parameters between agents, particularly in regards to farm success (\cite{miller}, \cite{valdivia},\cite{walsh}). In order to create a more detailed and perhaps more accurate simulation, the relationships between different factors on the island must be better understood. This work aims to search through a large number of possible relationships and identify the empirically most significant ones for future study and incorporation into simulation models.

@@

The data available to study the dynamics between agricultural measures and related factors primarily 
comes from the Censo de las Unidades de Producci\'on Agropecuaria (upa) de Gal\'apagos (Census of Agricultural Production 
Units (UPA) of Galapagos) (\cite{census}). This is a self-reported survey with data from 755 farms (UPAs) detailing many characteristics. The questions covered fall into the following categories:
\begin{itemize}
\item{General characteristics (land area, age of landowner, etc.)}
\item{Permanent crops (specific types and quantity)}
\item{Temporary crops (specific types and quantity)}
\item{Pastures (specific types and quantity)}
\item{Tree crops (specific types and quantity)}
\item{Livestock production (detailed by animal)}
\item{Expenses (detailed by category)}
\item{Workers (detailed by role)}
\item{Land use (8 different categories)}

\end{itemize}

In addition to this census, data is also available from satellite image classification including 
information on water, energy and road access. In total, under the direction of the client 239 variables were selected for 
consideration in modeling relationships between potential predictors and five outcome variables of interest.

Some of the five outcome variables of interest came directly from survey responses, while others were derived from a 
combination of multiple variables. When analyzing a derived outcome variable all variables used in its original calculation were 
removed from consideration in the model . The client also denoted specific predictor variables to exclude from particular 
models when their inclusion would not be beneficial. For example, while the 
amount of crops sold in pounds was not directly used to calculate net income, the obvious relationship existent precluded it 
from inclusion. In addition, predictor variables that met one or more of the following criteria were removed prior to modeling: 
zero variance, extremely high (>0.99) or perfect correlation with other predictor variables, or linear dependence with other 
predictor variables (that is, two or more predictor variables could be linearly combined to create another predictor variable). The 
exact number of predictor variables included in each model varied slightly, but there were approximately 200 predictors variables examined for each model after the preceding steps were taken.

@@

The remainder of this report is divided into four sections, Section 2: Modeling, Section 3: Results, Section 4: Statistical 
Methods, and Section 5: Limitations and Future Work. Section 2 provides a brief overview of the analysis so that the results can be understood. In Section 3 results for each of the 5 outcome variables are provided, where standard tables and graphs are repeated for each. A more in-depth explanation of the statistical methods used is contained in Section 4, but this section can be referenced as needed. Section 5 details important considerations when interpreting the results and suggests possible avenues for future work. Lastly, References and the Appendix, including additional tables and figures, can be found at the end.

@@

The primary challenge in this analysis is the vast number of potential predictor variables. This challenge is twofold; 1.) when 
the number of predictors is large the determination of a reliable model is difficult and 2.) interpreting the coefficients of many 
predictors simultaneously is not an easy task for humans (and will make resulting simulations overly complicated). For this 
reason, the analysis focuses on the use of two variable selection techniques that aim to build a linear model with a subset of 
the available variables that still maintains a strong explanatory/predictive performance.

Secondarily, when performing standard linear regression the error is assumed to be normally distributed.  While this does not 
mean the outcome variable necessarily needs to be normally distributed, large deviations from normality can cause issues. 
Several outcome variables in this study show strong non-normality, which can contribute to a poorly fitting model and unreliable estimates of the coefficients. In order to address this issue transformations to the data and modifications to the standard linear model will be considered.

@@

A summary of the statistical methods used is presented here to allow for understanding of results. For further details see 
Section 4: Statistical Methods. For each of the outcome variables of interest a set of linear models using the appropriate subset 
of predictors was built. Each relationship was modeled 
using Forward Selection and ElasticNet regression. Forward Selection fits a linear model by progressively adding variables to 
the model until a best fit is found. This results in only some of the variables being included, chosen in a discrete manner (computed using the R package `leaps' \cite{lumley}). 
ElasticNet regression fits a linear model by limiting the size of the coefficients so that they are smaller than in standard least 
squares, and for many variables actually shrunken to zero. Similar to Forward Selection this results in a smaller model, but 
variable selection can be carefully tuned as optimization is done in a more continuous way (computed using the R package `glmnet' \cite{friedman}).

In general, these techniques have slightly different aims. Forward Selection chooses a model that best explains the variance in 
the dataset at hand. Elasticnet chooses a model that can best make predictions on new data. Depending on the goals of 
analysis, one technique is not necessarily better so we do not compare the two quantitatively, but instead offer both results as 
varying perspectives on variable selection. While a variable being chosen by both methods provides stronger evidence that an 
important relationship exists, disagreement suggests exploring both possibilities instead of one method necessarily being 
incorrect.

@@

The first three outcome variables of interest can be grouped together under the category of farm success; labeled as 
productivity, net income and number of workers supported. Productivity is calculated as the total pounds of crops and livestock 
produced divided by the farm surface area. Net income is calculated as the the revenue from all products 
sold minus total expenses. Number of workers supported is calculated as the total labor expenditures divided by a standard 
full-time worker's salary.

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.8\textwidth]{production_histograms.pdf}
\caption{Histogram of Productivity, from left to right showing the full range, between 0-10,000 lbs/hectare, and the log transformed nonzero values. The log transformed values exhibit the desired normal shape.}
\label{figure:prod_hist}
\end{figure}

The histograms of the Productivity variable (Figure \ref{figure:prod_hist}) show a strong skewness, both when looking at all 
observations, and when zooming into observations between 0-10,000 lbs/hectare. To achieve a distribution closer to normal 
(bell-curved), which will benefit the linear model, we took the $log_{10}$ transformation with resulting data shown in the last 
plot. Since the log transformation cannot be performed on zero values, we removes the 38 occasions of this from the dataset. 
Beyond the mathematical constraint, farms with zero production perhaps are not farms as typically defined.

\begin{table}[h!]
\centering
\begin{tabular}{cc}
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection} 
\csvreader[head to column names]{production_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{Modeling of Productivity, Top 5 features for both methods}
\label{table:prod_top5}
\end{table}

We built linear models using both methods, Elasticnet and Forward Selection, on the log-transformed 
productivity variable, recording an optimal model of any size and the best 5 variable model for each. The size of 5 variables is 
chosen to provide quick insight and not for any specific statistical property. The results of the best 5 variable model are shown 
in Table \ref{table:prod_top5}, listed in alphabetical order.  Next to variable names the direction of the relationship is indicated 
with a (+) or (-). There is some overlap between the variables selected by each of the methods, but also unique choices made 
by each method. The Root Mean-Squared Error (RMSE) for the 5 variable Elasticnet model is 0.78 and the $R^2$ for 
Forward Selection is 0.48. The RMSE is on average how far the predicted value is from the actual value and $R^2$ is the 
percentage of variability in the outcome variable that is explained by the model. To provide context for the RMSE, we can look 
at the point furthest to the right on the cross-validation plot and see how the model would perform when no variables are 
included in the model, that is just predicting the average outcome value.

Plots from the optimal models for Elasticnet and Forward Selection are shown in Figure \ref{figure:prod_opt}. The cross-
validation plot for Elasticnet can be understood as follows: the 
horizontal axis shows the number of variables included in the model (on top) and the corresponding lambda ($\lambda$) value 
(on the bottom), the vertical axis shows the Mean-Squared Error (MSE) represented as the red dots and surrounded by bars 
showing the standard deviation. The vertical dashed line to the left, $\lambda_{min}$, is found at the minimum MSE and the 
vertical dashed line to the right, $\lambda_{1se}$ is at the largest lambda within one standard error of the minimum. The idea 
behind $\lambda_{1se}$ is that similar error performance can be achieved with a smaller model, in this case a model with 40 
fewer variables. Since our goal is to select a small amount of variables, we will generally use the model found at $
\lambda_{1se}$. For Forward Selection the plot is much more straightforward, we plot the number of variables included versus 
the information criterion that we choose to minimize and mark the optimal point in red.

\begin{figure}[h]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{elastic_cv_production.pdf}
\caption{Cross-validation plot for Elasticnet}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{forward_nvars_production.pdf}
\caption{Forward Selection plot}
\end{subfigure}
\caption{Productivity Variable Selection. The Elasticnet plot shows a desirable "U" shape where a moderate number of variables provides a significant improvement in prediction error. The optimal Forward Selection model includes 26 variables.}
\label{figure:prod_opt}
\end{figure}

The convex shape of the plots highlights a common trend in variable selection; not including enough variables does not provide 
enough information, but beyond a certain point adding more variables may not be worth the added complexity. The optimal 
model chosen for Elasticnet includes 45 variables (not counting the intercept term) and using Forward Selection we choose a 
model of 26 variables, almost all of which are also included in the Elasticnet model. The coefficients estimated for both models 
can be found in Table \ref{table:prod_full} in the Appendix. For the full models, the RMSE and $R^2$ are 0.61 and 0.63, 
respectively. These numbers suggest that while the 5 variable model is helpful, there is a decent amount of information to be 
gained by adding more variables. Diagnostics of the linear fit of the optimal Elasticnet and Forward Selection models (plots shown in Figure \ref{figure:resids_prod} in the Appendix) do not raise any concerns.

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.75\textwidth]{netincome_histograms.pdf}
\caption{Histograms of Net Income, with full data and data within one standard deviation displayed. The values very far from the center are of concern when modeling.}
\label{figure:netincome_hist}
\end{figure}

The histograms for Net Income (Figure \ref{figure:netincome_hist}) show a symmetric shape, but a very spiky center and a few 
observations wide in the tails. We attempted to fit a model on the full dataset, but find the observations with large absolute 
values are obscuring other possible information in the model. After trying several cutoff thresholds and examining model fit, we  removed the 28 observations that are beyond one standard deviation (30,478) from the mean in either direction. The 
following models are fit on this reduced dataset of 727 observations.

\begin{table}[h!]
\centering
\begin{tabular}{cc}
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{netincome_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{Modeling of Net Income, Top 5 features for both methods}
\label{table:netincome_top5}
\end{table}

The results of the best 5 variable model are shown in Table \ref{table:netincome_top5}, with a RMSE of 5790.35 for Elasticnet and a $R^2$ of 0.15 for Forward Selection. For interpretation of the RMSE it is important to keep in mind the scale for Net Income is much larger than that of log productivity. However, in this case adding variables has caused a minimal decrease in the error.
\begin{figure}[!tbp]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{elastic_cv_netincome.pdf}
\caption{Cross-validation plot for Elasticnet...}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{forward_nvars_netincome.pdf}
\caption{Forward Selection plot}
\end{subfigure}
\caption{Net Income Variable Selection. The Elasticnet plot shows a much smaller improvement in prediction error and Forward Selection includes a small number of variables.}
\label{figure:netincome_opt}
\end{figure}

The cross-validation plot for Net Income show wider error bars throughout the range of model sizes and just using the average 
Net Income would predict nearly as well as any other model. Since here $\lambda_{1se}$ only includes the intercept, we 
chose the optimal Elasticnet model to be at $\lambda_{min}$, which includes 9 variables. For Forward Selection, we also 
included 9 variables, with all but two overlapping with the Elasticnet choices. The coefficents estimated for both models can be 
found in Table \ref{table:netincome_full} in the Appendix. The full models had an RMSE and $R^2$ of 5768.30 and 0.19, 
respectively. Since the full models are not much larger than the 5 variable models, the small improvements are not surprising. Diagnostics of the linear fit of the optimal Elasticnet and Forward Selection models (plots shown in Figure 
\ref{figure:resids_netincome} in the Appendix) do not follow assumptions as closely as for the Production models, but are not so 
concerning as to disqualify the results. On the whole, the results from the various plots and diagnostics suggest that the relationships found for Net Income are worth investigating, but that a linear relationship is not very strong.

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.75\textwidth]{worker_histograms.pdf}
\caption{Histograms of Workers, showing the proportion of zeroes, farms with nonzero workers, and the log transformed number of workers. The log transformation helps remove the skewness, but some bi-modality still remains.}
\label{figure:workers_hist}
\end{figure}

The first plot in Figure \ref{figure:workers_hist} shows that a large percentage of the farms are not able to support any workers. 
For this reason, we divided up the modeling task for this outcome variable. First, we used logistic regression to model the 
binary variable of whether a farm supports zero or more than zero workers. Secondly, we used linear regression to model the 
quantity of workers for just those 264 farms with a positive number of workers supported.

\begin{table}[h]
\centering
\begin{tabular}{cc}
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{workers_binary_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{Modeling of Binary Workers , Top 5 features for each method}
\label{table:workers_binary_top5}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{cc}
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{workers_nonzero_top5names.csv}{}%
{\\\elasticnet & \forward}%
\newline
\end{tabular}
\caption{Nonzero Workers}
\caption{Modeling of Nonzero Workers, Top 5 features for both methods}
\label{table:workers_nonzero_top5}
\end{table}

Sometimes Elasticnet will simultaneously choose to eliminate multiple variables at the same time, in this case there is not a 5 
variable model so we show the 6 variable model for Elasticnet. For logistic regression we can measure performance with the 
misclassification rate, which on average is 0.31 on the test set for Elasticnet and 0.26 on the full dataset for Forward Selection. 
These can be compared to the 0.35 misclassification rate we would achieve if we simply predicted the majority class, zero 
workers, every time. For the linear model 5 variable model the RMSE is 0.46 and the $R^2$  0.30. The low misclassification rate coupled with low $R^2$ show that there is a fairly clear divide between farms that can or cannot support workers, but the specific number of workers is much harder to predict. Several variables are found to be most helpful for both the logistic regression and linear regression models, but there is still a fair bit of difference.

\begin{figure}[h]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{elastic_cv_workers_binary.pdf}
\caption{Cross-validation plot for Elasticnet}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{forward_nvars_workers_binary.pdf}
\caption{Forward Selection plot}
\end{subfigure}
\caption{Workers Binary Variable Selection. The Elasticnet plot shows ideal predictive performance when including 5-25 variables. The computation for logistic regression Forward Selection stops at the optimal value so the rest of the plot is not included as before.}
\label{figure:workers_binary_opt}
\end{figure}

\begin{figure}[h]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{elastic_cv_workers_nonzero.pdf}
\caption{Cross-validation plot for Elasticnet}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{forward_nvars_workers_nonzero.pdf}
\caption{Forward Selection plot}
\end{subfigure}
\caption{Nonzero Workers Variable Selection. Both plots show that adding variables does little to improve performance, highlighting a lack of linear relationship.}
\label{figure:workers_nonzero_opt}
\end{figure}

For logistic regression Elasticnet chose a model of size 6 and Forward Selection chose a model of size 7. Since these 
models are not much larger the misclassification rates remained at 0.31 and 0.26, respectively. For linear regression, the full 
models for Elasticnet and Forward Selection had a RMSE and $R^2$ of 0.46 and 0.34, respectively. The Elasticnet full model 
had 5 variables and the Forward Selection full model had 8 variables.  In the Appendix, the coefficients estimated for logistic 
regression models can be found in Table \ref{table:workers_binary_full} and the coefficients estimated for linear regression 
models can be found in Table \ref{table:workers_nonzero_full}.

@@

The analysis of Invasive Species follows much the same pattern as for Number of Workers Supported. Again, we have a large 
number of zeros in the outcome variable, farms with zero percent of their land covered by invasive species. We first modeled 
this binary variable using logistic regression and then performed linear regression on the log transformed values for the farms 
that have a percent coverage greater than zero. As can be seen in the far right plot of Figure \ref{figure:invasive_hist} there are 
two very small values (0.005{\%} and 0.08{\%} prior to taking the log). They effect the fit of the model and are negligibly above 
zero so we removed them from the linear model, leaving 155 farms for modeling.

\begin{figure}[h!]
\centering
\includegraphics[width = 0.75\textwidth]{invasive_histograms.pdf}
\caption{Histogram of Invasive Species. To the left: the proportion of zeroes, in the middle: farms with greater than zero, and to the right: the log transformed percent of invasive species. The two very small values on the right plot are removed.}
\label{figure:invasive_hist}
\end{figure}
